Testing for neutrality of a trait from within and between species variations

### 2023-05-15 manuscript/main.tex
Added data from the Zoonomia project published in the Science SI.

\title{A neutrality index for quantitative traits from within and between variations leveraging sequences evolution}

@@

% Abstract (≤ 250 words)
%TC:endignore
\begin{abstract}
    To determine whether a trait is neutrally evolving or under selection, methods have been developed both at the population and at the phylogenetic scale.
    At the population scale, methods do not integrate the changes over long evolutionary time.
    Conversely, at the phylogenetic scale, current methods usually discards the variance in traits between individuals, leaving out valuable information.
    The main goal of this project is to combine within-species and between-species variation of quantitative traits to determine the evolutionary regime of a trait.
    We define a neutrality index for a quantitative trait, leveraging both data at the population and phylogenetic scale.
    Moreover, this index leverages sequence divergence and polymorphism to normalize the variation of trait between and within species.
    Our neutrality index equals to one for a neutral trait, suggesting that traits for which this relationship is not verified are putatively under a selective regime.
    If our neutrality index is lower than one for traits putatively under stabilizing selection, and conversely greater than one fors trait putatively under diversifying selection.
    Finally, we show that our test is not sensitive to the assumption that population sizes and mutation rates are constant across the phylogeny, and automatically adjust for it.
\end{abstract}

@@

For a given trait, its genetic architecture is defined by the number of loci encoding the trait ($\NbrLoci$) and the random effect on the random of a mutation ($a$).
New mutations are thus generating variance for such a trait.
We define the expected effect of a mutation on the variance of a trait as $\RateMut$, given as:
\begin{gather}
    \RateMut \defEqual \NbrLoci \Multiply \E \left[ a^2 \right]. \label{eq:rate-mut}
\end{gather}

>>>>

\title{Testing for neutrality of a trait from within and between species variations}

@@

% Abstract (≤ 250 words)
%TC:endignore
\begin{abstract}
    To determine whether a trait is neutrally evolving or under selection, methods have been developed both at the population and at the phylogenetic scale.
    At the population scale, methods do not integrate the changes over long evolutionary time.
    Conversely, at the phylogenetic scale, current methods usually discards the variance in traits between individuals, leaving out valuable information.
    The main goal of this project is to combine within-species and between-species variation of quantitative traits to determine the evolutionary regime of a trait.
    We define a neutrality index for a quantitative trait, leveraging both data at the population and phylogenetic scale.
    Moreover, this index leverages sequence divergence and polymorphism to normalize the variation of trait between and within species.
    Our neutrality index equals to one for a neutral trait, suggesting that traits for which this relationship is not verified are putatively under a selective regime.
    If our neutrality index is lower than one for traits putatively under stabilizing selection, and conversely greater than one for trait putatively under diversifying selection.
    Finally, we show that our test is not sensitive to the assumption that population sizes and mutation rates are constant across the phylogeny, and automatically adjust for it.
\end{abstract}

@@

For a given trait, its genetic architecture is defined by the number of loci encoding the trait ($\NbrLoci$) and the random effect of a mutation on the trait ($a$).
New mutations are thus generating variance for such a trait.
We define the expected effect of mutations on the variance of a trait as $\RateMut$, given as:
\begin{gather}
    \RateMut \defEqual \NbrLoci \Multiply \E \left[ a^2 \right]. \label{eq:rate-mut}
\end{gather}
